---
title: "z04 - Exploratory Data Analysis"
author: "Denis Lopes"
date: "29/10/2020"
output: 
  html_document:
  toc: true
---

# ggplot

**IMPORTAT**  
Consult for good referente of ggplot2
**ggplot2 chetsheet**

##basic
load data
The qplot doesnt work very well in Markdown
```{r qplot color}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data = mpg, color = drv)
```

##line smooth
```{r qplot smooth}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```


##histograms
```{r histograms}
qplot(hwy, data = mpg, fill = drv)
```

##Panels - Facets
```{r facets collumn}
qplot(displ, hwy, data = mpg, facets = .~drv)
```

##Panels - Facets
```{r facets row}
qplot( hwy, data = mpg, facets = drv~., binwidth = 2)
```


##Density Smooth
```{r density smooth}
qplot( hwy, data = mpg, geom = "density" , color = drv)
```



##Scatterplots

```{r basic}
 qplot( displ, hwy, data = mpg)
```




```{r shape}
 qplot( displ, hwy, data = mpg, shape = drv)
```


```{r color and shape}
 qplot( displ, hwy, data = mpg, shape = drv, color = drv)
```

## boxplot


```{r}
qplot(drv,hwy,data=mpg,geom="boxplot")
qplot(drv,hwy,data=mpg,geom="boxplot",color = manufacturer)
```




##with lm (linear regretion)
```{r with lm }
 qplot( displ, hwy, data = mpg, shape = drv, color = drv, geom = c("point","smooth"), method = "lm")
```

```{r with lm +geom_smooth}
 qplot( displ, hwy, data = mpg, shape = drv, color = drv) + geom_smooth( method = "lm")
qplot(displ, hwy, data = mpg, color=drv, geom = c("point", "smooth"))
```




```{r with lm +  plus facet}
 qplot( displ, hwy, data = mpg, shape = drv, color = drv, facets = .~drv) + geom_smooth( method = "lm")
```

##ggplo2 basic components 

Basic Components
* A **data frame**  
* **aesthetic mappings** mapped to color, size  
* **geoms** Geometric object like points, lines, shapes  
* **facets** for conditional plots  
* **stats** statistical transformations like binning, quantiles, smoothing.  
* **scales** scale an aesthetic maps uses (e.g. male = red, female = blue)  
* **coordinate system**  




## layers plot

```{r layers summary}
head(mpg)
```




```{r layers summary }
 g = qplot( displ, hwy, data = mpg, shape = drv, color = drv, facets = .~drv) 
g + geom_smooth( method = "lm") #auto print plot object without saving

p =g + geom_smooth( method = "lm") 
print(p)     #Explicitly save and print ggplot object

summary(g)  #summary of plot

```
## verificar diferença aes hipotese - data de g



```{r verificar}
#essa
g + geom_point(aes(color = drv), size = 4, alpha = 1/2)

#com essa
g + geom_point(color = drv, size = 4, alpha = 1/2)
```


##teste manipulations


```{r}
g = qplot( displ, hwy, data = mpg) 
g + geom_point() +
   geom_smooth() + # method = "lm") #teste 1 
  facet_grid(.~drv) 

```


## data lega

```{r}
qplot(carat,price,data=diamonds, shape=cut)
```





## annotation - legend, labers, transparent

```{r labels}
qplot( displ, hwy, data = mpg)  +
  geom_point(color = "red",size = 2, alpha = 1/2) +
  labs(title = "MPG data", x = expression("displ " * PM[2.45])) +
  geom_smooth(method = "lm", size = 1, linetype = 2, se=TRUE) +
  theme_bw()
  
```


## remove NA value

```{r}
qplot( displ, hwy, data = mpg, na.rm = TRUE)
```


## Cutpoints

```{r}
## Calculate the deciles of the data!
cutpoints <- quantile(mpg$displ, seq(0, 1, length = 5), na.rm = TRUE)


## Cut the data at the deciles and create a new factor variable!
mpg$displDecile <- cut(mpg$displ, cutpoints)
levels(mpg$displDecile)

qplot( displ, hwy, data = mpg)+ facet_grid(.~displDecile)  


```





##Becareful use axis limits , out line 

whem use ylim ... it cut of data out of plot  
for this case, use coord_cartesian(ylim = c(-3,3))


```{r use axis limits}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100 ## Outlier!
# plot small
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
# overall plot
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()
# outlier missing in x=50
g + geom_line() + ylim(-3, 3) + geom_rect(aes(xmin = 45, xmax = 55, ymin = -2, ymax = 2), fill = NA, colour = "red")
# outlier include
g + geom_line() + coord_cartesian(ylim = c(-3, 3))
```












